### 2021-08-27 manuscript/field_data.tex

The field data application was made with the aeromagnetic data of Carajás, Pará, Brazil, provided by the CPRM. 
The total number of data is 7,706,153. For this case we gridded the data to $500 \times 500 = 2,500,000$ observation points, which means that storing the full sensitivity matrix would be necessary 45.47 Terabytes of RAM. However, taking advantage that the second derivatives of equation \ref{eq:r} are symmetric or skew-symmetric matrices, it is possible to reconstruct the whole sensitivity matrix storing only the first column of each component of equation \ref{eq:Hi}, thus, using only 114.44 Megabytes.

To achieve high efficiency in property estimative of the equivalent sources, the method CGLS for inversion was used, combined with a fast matrix-vector product, only possible because of the Block-Toeplitz Toeplitz-Block(BTTB) structure of the sensitivity matrix. This fast matrix-vector product was also used for data processing (upward continuation) in a very efficient way.

As this area is very large different values of the magnetic main field can be considered. 
For this test it was considered an approximated mid location of the area (latitude $-6.5^{\circ}$ and longitude $-50.75^{\circ}$) where the declination is $-19.86^{\circ}$ for the IGRF model in 01/01/2014. The inclination was calculated considering the Geocentric axial dipole model ($tang \, I = 2 \times tan \, \lambda$) and is equal to $12.84^{\circ}$.
As the source magnetization is unknown, inclination and declination equals to the main field is being used.

Figure \ref{fig:carajas_real_data_mag} shows the observed magnetic field data of the area. Using a equivalent layer at $300$ meters above the ground the predicted data and its residual are shown in figure \ref{fig:carajas_gz_predito_mag}. The mean of $-2.7135\times 10^{-07}$ nT and the standart deviation of $3.2726\times 10^{-06}$ nT of the residual shows the good result of physical property estimative. It was used 50 iterations of the CGLS method taking $15.5$ seconds with a Intel core i7 7700HQ@2.8GHz processor.

In figure \ref{fig:up5000_carajas_500x500_mag} the upward continuation transformation was made at $5000$ meters and took $0.49$ seconds.

>>>>

The field data application was performed with the aeromagnetic data of Carajás, Pará, Brazil, provided by CPRM.
The survey is composed of $131$ flight lines N-S oriented, spacing $\Delta y = 3,000$ m. The magnetometer (Scintrex CS-3) was set to a interval between measurements of $0.1$ s giving a spacing $\Delta x = 7.65$ m. The average flight heigth is $\Delta z = -900$ m. The total number of observation points is $N = 6,081,345$. Figure \ref{fig:carajas_real_data_mag} shows the observed magnetic field data of the area.

For the actual data processing, we have made a comparison between an interpolated regular grid of $10,000 \times 131$ using a nearest neighbour algorithm and a decimated irregular grid, also of $10,000 \times 131$, totaling $N = 1,310,000$ observation points in both cases. The decimated grid was performed by using the regular grid created in the first case as a guide and by finding the nearest real observation point to this regular grid thus, ensuring that the irregular grid is the lesser deviant possible to conduct the BTTB scheme. In figure \ref{fig:carajas_real_data_decimated_gridline}a we show the result of the interpolation and in \ref{fig:carajas_real_data_decimated_gridline}b the result of the decimation. With $1,310,000$ observation points, it would be necessary $12.49$ Terabytes of RAM to store the full sensitivity matrix with the classical approach. However, taking advantage that the second derivatives of equation \ref{eq:r} are symmetric or skew-symmetric matrices, it is possible to reconstruct the whole sensitivity matrix storing only the first column of each component of equation \ref{eq:Hi}, thus, using only $59.97$ Megabytes, allowing desktop computers being able to process this amount of data.

As this area is very large, different values of the magnetic main field can be considered. 
For this processing, it was considered an approximated mid location of the area (latitude $-6.5^{\circ}$ and longitude $-50.75^{\circ}$) where the declination is $-19.86^{\circ}$ for the IGRF model in 1st january, 2014. The inclination was calculated considering the Geocentric axial dipole model ($tang \, I = 2 \times tan \, \lambda$) and is equal to $12.84^{\circ}$. As the source magnetization is unknown, inclination and declination equals to the main field is being used.

To achieve high efficiency in property estimative of the equivalent sources, the method CGLS for inversion was used, combined with a fast matrix-vector product, only possible because of the BTTB structure of the sensitivity matrix. This fast matrix-vector product was also used for data processing (upward-continuation) in a very efficient way.

Using a equivalent layer at $300$ meters above the ground the predicted data and its residual of the interpolated regular grid are shown in figure \ref{fig:carajas_gz_predito_mag_gridline}. The mean of $0.07979$ nT and the standart deviation of $0.5060$ nT of the residual shows the good result of physical property estimative. It was used 200 iterations of the CGLS method taking $390.80$ seconds with a Intel core i7 7700HQ@2.8GHz processor in single-processing and single-threading modes. Using the same equivalent layer and CGLS configuration the predicted data and its residual of the decimated irregular grid are shown in figure \ref{fig:carajas_gz_predito_mag_decimated}.
With a mean of $0.07348$ nT, standart deviation of $0.3172$ nT and lower residual amplitude compared to the regular grid, we show that the process of decimating the original data, without creating new observation points with interpolation, can be benefical to the method, even if an irregular grid is taken place. It took $385.56$ seconds to complete the estimative.

The convergence analysis for the decimated irregular grid up to $2,000$ CGLS iterations is in figure \ref{fig:convergence_carajas_mag_decimated}, showing good convergence rate and guaranteeing that the irregular grid is not disturbing the method.

In figure \ref{fig:up5000_carajas_decimated_mag} the upward-continuation transformation using the estimated equivalent layer with the decimated grid was made in a horizontal plane at $5,000$ meters and took $2.64$ seconds, showing good results without visible errors or border effects problems and accentuating the long wavelenghts.

### 2021-08-27 manuscript/introduction.tex

Equivalent layer background:

\cite{dampney1969}, \cite{blakely1996}, \cite{emilia1973}, \cite{li2014using}
\\\\
Fast methods for equivalent layer:

\cite{leao-silva1989}, \cite{mendonca-silva1994}, \cite{oliveirajr-etal2013}, \cite{li-oldenburg2010}, \cite{siqueira-etal2017}, \cite{mendoncca2020subspace}
\\\\
Methods for solving Toeplitz matrices:

\cite{golub-vanloan2013}, \cite{levinson1946}, \cite{chan-jin2007}
\\\\
Toeplitz matrices in Potential methods:

\cite{zhang-wong2015}, \citealp{zhang-etal2016}, \cite{hogue2020tutorial}, \cite{renaut2020fast}
\\\\
Convolutional equivalent layer grav:

\cite{siqueira-etal2017}, \cite{takahashi2020convolutional}
\\\\
About this work:

>>>>

Throughout the years many authors presented solutions to inverse problems in potential methods using magnetic data. The most common approaches are the works that try to parameterize the geometric source \citep{bott1960use,pilkington1986determination,bhattacharyya1980generalized} and others tried to estimate the susceptibility or the magnetization as the parameter \citep{parker1974inversion,cribb1976application,li19963,liu20173d}. With advances in computation power, the equivalent layer is an inversion technique that is becoming more present in data processing nowadays, despite its use since 1960 in geophysics literature \citep{danes1961structure,bott1967solution,dampney1969}. The equations deductions of the equivalent layer as a solution of the Laplace's equation in the region above the source was first presented by \cite{kellogg1929} and detailed explanations can also be found in \cite{blakely1996}. In magnetic data procesing, some authors explored this technique for calculating the first and second vertical derivatives fields \citep{emilia1973}, reduction to the pole \citep{silva1986,oliveirajr-etal2013,li2014using}, upward/downward continuations \citep{hansen-miyazaki1984,li-oldenburg2010} and total magnetic induction vector components calculation \citep{sun2019constrained}.

Together with the rise in computational processing, some works tried new implementations to increase the efficiency of the equivalent layer. In \cite{leao-silva1989} the authors used a shifting window over the layer, increasing the number of linear systems to be solved, but decreasing the size of it. In \cite{li-oldenburg2010} a method of transforming the full sensitivity matrix into a sparse one by using wavelets was another approach. \cite{oliveirajr-etal2013} also proposed a shifting window method, but instead of directly calculating the equivalent sources physical properties, they estimated the coefficients of a bivariate polynomial funtion representing the sources. An iterative method for gravimetric data was proposed by \cite{siqueira-etal2017}, where the mass distribution over the layer is corrected at each iteration by the residuals of observed and estimated data.

In \cite{takahashi2020convolutional}, the authors combined the fast equivalent source technique presented by \cite{siqueira-etal2017} with the concept of symmetric block-Toeplitz Toeplitz-block (BTTB) matrices to introduce the convolutional equivalent layer for gravimetric data technique. When the sensitivity matrix of the linear system, required to solve the gravimetric equivalent layer, is calculated on a regular spaced grid of dataset with constant height and each equivalent source is exactly beneath each observed data point, the BTTB structure appears. This work showed an decrease in the order of $10^4$ in floating-point operations needed to estimate the equivalent sources.

In potential field methods, the properties of Toeplitz system have been used for downward continuation \citep{zhang-etal2016} and for 3D gravity-data inversion using a 2D multilayer model \citep{zhang-wong2015}. More recently in \cite{hogue2020tutorial} the authors provided an overview on modeling the gravity and magnetic kernels using the BTTB structures and \cite{renaut2020fast} used for inversion of both gravity and magnetic data to recover sparse subsurface structures.

A wide variety of applications in mathematics and engineering that fall into Toeplitz systems propelled the development of a large variety of  methods for solving them. Direct methods were conceived by \cite{levinson1946} and by \cite{trench1964}. Currently, the iterative method of conjugate gradient is used in most cases, in \cite{chan-jin2007} the authors presented an introduction on the topic for 1D data structures of Toeplitz matrices and also for 2D data structures, which they called block-Toeplitz Toeplitz-block matrices. In both cases, the solving strategy is to embbed the Toeplitz/BTTB matrix into a Circulant/Block-Circulant Circulant-Block matrix, calculate its eigenvalues by a 1D or 2D fast Fourier transform of its first column, respectively and carry the matrix-vector product between kernel and parameters at each iteration of the conjugate gradient method in a very fast manner.

In this work, the convolutional equivalent layer using the block-Toeplitz Toeplitz-block idea, presented in \cite{takahashi2020convolutional}, will be used to solve the linear system required to estimate the physical property that produces a magnetic field on regular grids. Despite the non-existence of a fast equivalent source technique solver as with gravity data, it is possible to achive very fast solutions using a conjugate gradient algorithm combined with the fast Fourier transform. We present a novel method of exploring the symmetric structures of the second derivatives of the inverse of the distance contained in the magnetic kernel, to keep the memory RAM usage to the minimal by using only one equivalent source to carry the calculations of the forward problem. We also show tests of the magnetic convolutional equivalent layer when irregular grids are used. The convergence of the conjugate gradient maintains in an acceptable level even under strong deviations of the data acquisition line. This shows that a fast and robust technique is possible when processing large amounts of magnetic data using the equivalent layer.
